Functional assembly into and among community of a tropical evergreen forest from Western Ghats, India

### 2016-08-04 inst/doc/draft.Rmd
v0.1.9027 LDMC ~ Canopy

---
title: 'Seasonal filtering at community scale of a tropical evergreen forest from Western Ghats, India'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(parallel)
library(Uppangala)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

@@

Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction. Functional traits can thus be linked to trade-offs representative of species ecological strategies. For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary in so called economic spectra. @wright_worldwide_2004 showed the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009. The leaf economic spectrum opposes productive ecological strategy to conservative persistent species. The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. A debated question is links between species economic spectra and strategy axes. Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.

@@

<!-- Tropical evergreen forest are composed by tree keeping their foliage all year around. Yet species will observe leaves fall events ruled by their leaf lifespan. Leaf lifespan will tend to decrease in tropical evergreen forest seasonally dry [@poorter_leaf_2009; @prior_leaf_2003]. -->

+ *Move to India Western Ghats context*
+ *No study of community ecology with functional traits*
+ *Peculiar case of highly seasonal moist evergreen forest*
+ *Move to previous questionning*
+ *Quid from ecological strategies at small scale in Uppangala forest*
+ *Quid from partitionning between stochastic event and environmental filtering on community structure*

@@

PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
PFT_sp$ID <- row.names(PFT_sp)
PFT_sp <- merge(PFT_sp, PCAscores, all = T)
row.names(PFT_sp) <- PFT_sp$ID
PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
row.names(Q) <- PFT_sp$SpCode
rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
```

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
sel <- intersect(row.names(Q), names(L))
Q <- Q[sel,]
L <- L[sel]
R <- R[row.names(L),]
L <- list(abundance = L, `presence-absence` = L)
L$`presence-absence`[L$`presence-absence` > 1]  <- 1
rm(Deriv, XY, com, Env, Trees, sel, BA)
```

@@

# Simulated CWMs
n = 999
nQ <- rep(list(Q), n)
nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
cl <- makeCluster(3)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterApply(cl, L, function(l){
  lapply(nQ, function(q){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
      apply(l, 1, function(x){
        weighted.mean(q[,v], x, na.rm=T)
      })
    })
  })
})
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
## Regressions
nCWM.lm <- mapply(function(x,y){
  lapply(y, function(z){
    mapply(function(r,s){
      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
    }, r = x, s = z, SIMPLIFY = F)
  })
}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
rm(nCWM) ; invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
nFval <- lapply(nFval, function(f){lapply(f, unlist)})
pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
row.names(pval) <- row.names(Fval)
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
ppval$abundance$BA <- NA
ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
ppval <- lapply(ppval, function(x){1 - x})
rm(pFval, npFval)

@@

cl <- makeCluster(3)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterApply(cl, nL, function(l){
  lapply(as.list(names(CWM$abundance)), function(v){
    apply(l, 1, function(x){
      weighted.mean(Q[,v], x, na.rm=T)
    })
  })
})
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
nCWV <- mapply(function(l,cwm){
  lapply(names(CWM$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
rm(nCWM.cwv, cl, nL) ; invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
nCWV <- lapply(nCWV, data.frame)
CWV.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.pval.res <- data.frame(inf = apply(apply(CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}))
## CWV SES
CWV.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
CWV.ses.res <- lapply(CWV.ses.res, data.frame)
CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CWV.ses.res <- CWV.ses.res[-1]
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
CWV.ses.res[,setdiff(var, names(CWV.ses.res))] <- NA
CWV.ses.res <- CWV.ses.res[var]
CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
row.names(CWV.ses.res) <- names(CWV.ses)
CWV.ses.res[is.na(CWV.ses.res)] <- ''
CWV.ses.res <- data.frame(CWV.ses.res)
CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)

@@

# Simulated CVNNDs 
cl <- makeCluster(3)
clusterExport(cl, list('nQ', 'L'))
nCVNND <- clusterApply(cl, nQ, function(x){
  apply(L$abundance, 1, function(l){
    apply(x, 2, function(q){
      cati::CVNND(rep(q,l), na.rm = T)
    })
  })
})
stopCluster(cl) ; rm(cl)
rm(nQ) ; invisible(gc)
## Non null CVNNDs
nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
nCVNND <- lapply(nCVNND, data.frame)
CVNND.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.pval) <- colnames(CVNND)
CVNND.pval <- apply(CVNND.pval, 2, p.adjust, method = 'fdr')
CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
## CVNND SES
CVNND.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.ses) <- colnames(CVNND)
CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CVNND.ses.res <- CVNND.ses.res[-1]
CVNND.ses.res[,setdiff(var, names(CVNND.ses.res))] <- NA
CVNND.ses.res <- CVNND.ses.res[var]
CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
row.names(CVNND.ses.res) <- names(CVNND.ses)
CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
CVNND.ses.res <- data.frame(CVNND.ses.res)
CVNND.ses.res$`$R^2$` <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)

@@

# Simulated mantel
nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
cl <- makeCluster(3)
clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
nspaflo <- lapply(nL$perm, function(l){spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'})
nspaflo.mantel <- lapply(nspaflo, function(ns){
  list(`E` = vegan::mantel(ns, Me, permutations = 0)$statistic, 
       `S | E`= vegan::mantel.partial(ns, Ms, Me, permutations = 0)$statistic)})
stopCluster(cl)
rm(cl, nL) ; invisible(gc())
## Non null mantels
nspaflo.mantel <- lapply(nspaflo.mantel, `[`, names(nspaflo.mantel[[1]]))
nspaflo.mantel <- apply(do.call(rbind, nspaflo.mantel), 2, as.list)
nspaflo.mantel <- lapply(nspaflo.mantel, unlist)
spaflo.res <- mapply(function(s, ns){
    paste(format(s, scientific = T, digits = 3),
          Uppangala::stars(min(1-rank(c(s,ns))[1]/(n+1),rank(c(s,ns))[1]/(n+1)), ns = '   '))
}, s = spaflo.mantel, ns = nspaflo.mantel)
spaflo.res <- data.frame(spaflo.res)
names(spaflo.res) <- c('$I_{st}$')

@@

The leaf economic specrtum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC. Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006. @bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds. Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species. Curvature is also affecting the leaf economy. Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC. SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011]. **LDMC rare species <- Canopy height**

>>>>

---
title: Functional assembly into and among community of a tropical evergreen forest from Western Ghats, India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(parallel)
cores <- detectCores() - 1
library(Uppangala)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

@@

Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction. Functional traits can thus be linked to trade-offs representative of species ecological strategies. For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary in so called economic spectra. @wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009. The leaf economic spectrum opposes productive ecological strategy to conservative persistent species. The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. A debated question is links between species economic spectra and strategy axes. Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.

@@

+ *Move to India Western Ghats context*
    - *No previous study of community ecology with functional traits*
    - *Peculiar case of highly seasonal moist evergreen forest*
+ *Move to previous questionning*
+ *Quid from ecological strategies at small scale in Uppangala forest*
+ *Quid from partitionning between stochastic event and environmental filtering on community structure*

@@

PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
PFT_sp$ID <- row.names(PFT_sp)
PFT_sp <- merge(PFT_sp, PCAscores, all = T)
row.names(PFT_sp) <- PFT_sp$ID
PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
```

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
row.names(Q) <- PFT_sp$SpCode
sel <- intersect(row.names(Q), names(L))
Q <- Q[sel,]
L <- L[sel]
R <- R[row.names(L),]
L <- list(abundance = L, `presence-absence` = L)
L$`presence-absence`[L$`presence-absence` > 1]  <- 1
rm(Deriv, XY, com, Env, Trees, sel, BA)
```

@@

# Simulated CWMs
n = 999
nQ <- rep(list(Q), n)
nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
cl <- makeCluster(cores)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterApply(cl, L, function(l){
  lapply(nQ, function(q){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
      apply(l, 1, function(x){
        weighted.mean(q[,v], x, na.rm=T)
      })
    })
  })
})
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
## Regressions
nCWM.lm <- mapply(function(x,y){
  lapply(y, function(z){
    mapply(function(r,s){
      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
    }, r = x, s = z, SIMPLIFY = F)
  })
}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
rm(nCWM) ; invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
nFval <- lapply(nFval, function(f){lapply(f, unlist)})
pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
row.names(pval) <- row.names(Fval)
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
ppval$abundance$BA <- NA
ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
ppval <- lapply(ppval, function(x){1 - x})
rm(pFval, npFval)

@@

cl <- makeCluster(cores)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterApply(cl, nL, function(l){
  lapply(as.list(names(CWM$abundance)), function(v){
    apply(l, 1, function(x){
      weighted.mean(Q[,v], x, na.rm=T)
    })
  })
})
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
nCWV <- mapply(function(l,cwm){
  lapply(names(CWM$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
rm(nCWM.cwv, cl, nL) ; invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
nCWV <- lapply(nCWV, data.frame)
CWV.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.pval.res <- data.frame(inf = apply(apply(CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}))
## CWV SES
CWV.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
CWV.ses.res <- lapply(CWV.ses.res, data.frame)
CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CWV.ses.res <- CWV.ses.res[-1]
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
CWV.ses.res[,setdiff(var, names(CWV.ses.res))] <- NA
CWV.ses.res <- CWV.ses.res[var]
CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
row.names(CWV.ses.res) <- names(CWV.ses)
CWV.ses.res[is.na(CWV.ses.res)] <- ''
CWV.ses.res <- data.frame(CWV.ses.res)
CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)

@@

# Simulated CVNNDs 
cl <- makeCluster(cores)
clusterExport(cl, list('nQ', 'L'))
nCVNND <- clusterApply(cl, nQ, function(x){
  apply(L$abundance, 1, function(l){
    apply(x, 2, function(q){
      cati::CVNND(rep(q,l), na.rm = T)
    })
  })
})
stopCluster(cl) ; rm(cl)
rm(nQ) ; invisible(gc)
## Non null CVNNDs
nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
nCVNND <- lapply(nCVNND, data.frame)
CVNND.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.pval) <- colnames(CVNND)
CVNND.pval <- apply(CVNND.pval, 2, p.adjust, method = 'fdr')
CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
## CVNND SES
CVNND.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.ses) <- colnames(CVNND)
CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CVNND.ses.res <- CVNND.ses.res[-1]
CVNND.ses.res[,setdiff(var, names(CVNND.ses.res))] <- NA
CVNND.ses.res <- CVNND.ses.res[var]
CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
row.names(CVNND.ses.res) <- names(CVNND.ses)
CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
CVNND.ses.res <- data.frame(CVNND.ses.res)
CVNND.ses.res$`$R^2$` <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)

@@

# Simulated mantel
nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
cl <- makeCluster(cores)
clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
nspaflo <- lapply(nL$perm, function(l){spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'})
nspaflo.mantel <- lapply(nspaflo, function(ns){
  list(`E` = vegan::mantel(ns, Me, permutations = 0)$statistic, 
       `S | E`= vegan::mantel.partial(ns, Ms, Me, permutations = 0)$statistic)})
stopCluster(cl)
rm(cl, nL) ; invisible(gc())
## Non null mantels
nspaflo.mantel <- lapply(nspaflo.mantel, `[`, names(nspaflo.mantel[[1]]))
nspaflo.mantel <- apply(do.call(rbind, nspaflo.mantel), 2, as.list)
nspaflo.mantel <- lapply(nspaflo.mantel, unlist)
spaflo.res <- mapply(function(s, ns){
    paste(format(s, scientific = T, digits = 3),
          Uppangala::stars(min(1-rank(c(s,ns))[1]/(n+1),rank(c(s,ns))[1]/(n+1)), ns = '   '))
}, s = spaflo.mantel, ns = nspaflo.mantel)
spaflo.res <- data.frame(spaflo.res)
names(spaflo.res) <- c('$I_{st}$')

@@

The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC. Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006. @bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds. Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species [@soman_aspects_1990]. Curvature is also affecting the leaf economy. Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC. SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].

Biotic factors also affect the LES through the negative effect of canopy height on LDMC. The canopy height effect in fact is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown). We could have assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010]. But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious. Instead of a growth-survival trade-off shade tolerant species may adopt a more low-high light growth trade-off strategy [@poorter_leaf_2009] strongly affecting SLA but not driving LDMC. Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance. Thus canopy species with the lowest LDMC will grow the fastest and then reach the highest canopy. Measurement of LDMC on species classified as canopy and emergent are realized most of the time on shadow leaves from understorey individuals. Still we observed previous relation indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times bigger than intra-individuals variation.
